Synthesising and Verifying Closed-loop Digital Control Systems

### 2016-09-23 paper/paper.tex
cleanout
crlf -> lf
line breaks

% This is "sig-alternate.tex" V2.1 April 2013
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% -------------------------------------------------------------------------------------------------% For tracking purposes - this is V2.0 - May 2012

@@

First, some assumptions have to be established to obtain a model for $\Delta{G_{Q}(z)}$.
%
\begin{myassumption}
\label{whitenoise}
The quantization noises $\nu_{1}$ (from the $Q1$) and $\nu_{2}$ (from the $Q2$) are white noises with zero mean.
\end{myassumption}

@@

\begin{myassumption}
The DAC and ADC are synchronized, {\it i.e.}, there is no delay between sampling the plant output at ADC and updating the DAC accordingly.
\end{myassumption}

\begin{myassumption}
The controller and plant are linear. In particular, the digital controller does not contain nonlinearities such as overflows, saturation, and limit-cycle oscillations.
\end{myassumption}

\begin{myassumption}
The DAC interpolator is an ideal ZOH; it takes the discrete value of control signal $u(k)$ at instant $k$ and produces a rectangular pulse of height $u(k)$ and width equal to the sample time $T$.
\end{myassumption}

Given those assumptions, the present approach is further based on the following lemmas.
%
\begin{mylemma}
\cite{widrow1956} The second moment (variance) of the PDF of quantization noise is given by \textcolor{red}{expand the acronym PDF}
\begin{equation}
\label{eq:variancelemma}
\sigma_{Q}^{2}=\frac{q^{2}}{12},
\end{equation}
\end{mylemma}

@@

\begin{mylemma} \cite{astrom1997computer}
The pulse transfer function $G(z)$ is equivalent at the sample times to the continuous transfer function $G(s)$ driven by ZOH. This pulse transfer function can be computed as
\begin{equation}
\label{eq:pulsetf}
G(z)=(1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}
\end{mylemma}

@@

\noindent where $\sigma_{Q}^{2}$ is the variance of quantization noise and $U(z)$ is the z-transform of control signal ({\it i.e.}, the z-transform of the signal from the digital controller), which is provided to the plant by means of ZOH.
%
\begin{proof}
Suppose that the plant discrete model in Figure~\ref{fig:sampledsystem} can be represented by the following pulse transfer function (without quantization noise effect): 
\begin{equation}
\label{eq:tfwithout}
G(z)=\frac{Y(z)}{U(z)}=\frac{B(z)}{A(z)}=\frac{b_{0}+b_{1}z^{-1}+b_{2}z^{-2}+\cdots+b_{M}z^{-M}}{1+a_{1}z^{-1}+a_{2}z^{-2}+\cdots+a_{N}z^{-N}},
\end{equation}
\noindent where $Y(z)$ is the z-transform of the plant output signal without noise $y(k)$, $U(z)$ is the z-transform of the control signal, which is provided to the plant, $A(z)$ and $B(z)$ are the plant pulse transfer function numerator and denominator, respectively.

@@

Eq.~\eqref{eq:deltag_var} indicates that the modeling of quantization noise effect on the plant transfer function demands an estimation of the variance of quantization noise $\sigma_{Q}^{2}$. To obtain such estimation, the methodology presented by Widrow and Koll\'ar~\cite{widrow2008quantization} is used. The authors provide an algorithm for computing the mean square of the output noise due to $Q1$ and $Q2$. Considering that each quantizer contribution can be represented by a white noise with zero mean (Assumption~\ref{whitenoise}), the mean square can be considered as the variance of noise ($\sigma_{Q}^{2}$), which can be calculated as
%
\begin{equation}
\sigma_{Q}^{2}=\frac{\lambda_{1}^{2}\cdot \sum {h_{Q1}^{2}}+\lambda_{2}^{2}\cdot \sum {h_{Q2}^{2}}}{12},
\end{equation}

\noindent where $\lambda_{1}$ and $\lambda_{2}$ are the scale factor for the quantization $Q1$ and $Q2$, respectively; and $\sum {h_{Q1}^{2}}$ and $\sum {h_{Q2}^{2}}$ are the squares sums of the impulse response obtained by means of a transfer function from $Q1$ and $Q2$ to the plant output, respectively.

@@

These equations are sufficient for computing $\sigma_{Q}^{2}$. However, there is still a gap for a complete model for $\Delta{G_{Q}(z)}$ in Eq.~\eqref{eq:deltag_var}, {\it i.e.}, the z-transform of control action $U(z)$. This indicates that the quantization noise effect on a discrete plant model mainly depends on three main factors:
%
\begin{itemize}
\item \textbf{Sample time:} the sample time effect is reflected over pulse transfer function that covers the effects of DAC ant its interpolator (ZOH);
\item \textbf{ADC and DAC step: } these are the main component of quantization noise power, but an $8$-bits converter already reduces the effect of quantization noises in the output for negligible values.
\item \textbf{Signal-to-noise ratio (SNR): } the dependence on the value of $U(z)$ shows the importance of the SNR for handling quantization noise; for a high SNR of the digital controller, the quantization effect is negligible. This happens during the transient of the system, when high control actions are employed. However, it can be relevant during the steady state. A digital controller with integral action might be sufficient for eliminating such influence on the steady state.
\end{itemize}

@@

\noindent where $R(z)$ can be considered as a step signal (for stabilization purpose) with height equal to $r$, which can be rewritten as
\begin{equation}
\Delta{G_{Q}(z)}=\frac{\sigma^{2}_{Q}}{r}\cdot \frac{1+C(z)G(z)}{C(z)} \cdot \frac{z-1}{z}.
\end{equation}

@@

The problem of parametric control synthesis based on stability measures for continous Linear Time Invariant (LTI) Single Input-Single Output (SISO) systems has been researched for several decades. On a theoretical level it is a solved problem\cite{wonham1967pole} for which researches continuously seek better results based on a number of properties in addition to stability. A vast range of pole placement techniques such as Moore's algorithm for eigenstructure assignment \cite{klein1977eigenvalue} or the more recent Linear Quadratic Regulator (LQR) \cite{bemporad2002explicit} have been used in several papers with increasing degrees of succes to ensure stability. The latter approach highlights the importance of conserving energy during the control process, which results amongst other things in lower running costs. Since real systems are subject to tolerance and noise as well as the need for economy, more recent works focus on the problems of achieving robust stability with minimum gain \cite{schmid2014unified,konigorski2012pole}. 
However, when extending the problem to digital controller synthesis, many of these techniques lack the ability to produce sound or stable results because they disregard the effects of quantization and rounding. Modern papers on implementations/synthesis of LTI digital controllers \cite{das2013lqr,ghosh2013fpga} focus on time discretization, failing to account for these error-inducing effects and can therefore result in digitally unstable systems that have been proven to be robustly stable in a continuous space. This situation demands the inclusion of formal methods to verify the effects of space discretization on the implementation of the controller.

@@

Researchers in formal verification methods of control systems have studied various effects of discretising the dynamics, including delayed response \cite{Duggirala2015}, and Finite Word Length (FWL) constraints with the objective to either verify (eg. \cite{daes20161}) or optimise (eg. \cite{oudjida2014design}) implementations. There are in fact two different problems arising from FWL effects. The first corresponds to the error in the dynamics caused by the inability of the hardware to represent the exact dynamics of the system whilst the second relates to rounding errors during computation. 
In \cite{fialho1994stability} a stability measure based on the error of the digital dynamics ensures that the FWL errors do not make the digital system unstable. A more recent approach \cite{DBLP:journals/automatica/WuLCC09} uses $\mu$ calculus to directly model the digital controller so that the selected parameters are stable by design. Most papers following this line of verification focus on finding a correct implementation of a known desired controller, looking for optimal parameter representations using FWL but ignoring the effects of rounding errors due to issues of mathematical tractability.
The analyses in \cite{DBLP:conf/hybrid/WangGRJF16,DBLP:conf/hybrid/RouxJG15} rely on an invariant computation on the discrete system dynamics using Semi-Definite Pogramming (SDP). Whilst the former uses BIBO properties to determine stability, the latter uses Lyapunov-based quadratic invariants. In both cases the SDP solver uses floating point arithmetic and soundness is checked by upper bounding the error.

@@

To the authors' knowledge, there is no existing work on the automatic synthesis of digital controllers considering FWL effects. Parameter synthesis tools such as \cite{cimatti2013parameter} use bounded model checking and fixed point computations to find parameters based on user defined specifications, but they often operate in the continuous domain and are ill-suited for robust analysis since they have no direct means of evaluating robustness. Other tools such as \cite{economakos2016automated} are directly aimed at robust stability problems but fail to explore the effects of FWL in their implementation.
In order to provide a correct by design controller, \cite{alur2016compositional} requires a user-defined finite state abstraction to synthesise a digital controller based on high level specifications. Whilst this approach directly overcomes the challenges presented by the FWL problem, it still requires error-prone user intervention. A different solution that looks at the other end of the scale is an approach that synthesises word lengths for know problems as developed in \cite{jha2013swati}, however this provides neither an optimal word synthesis nor a comprehensive solution for the problem.
It is from this vacuum that we seek to explore a CEGIS approach to FWL controller synthesis.

@@

In recent years the development of program synthesis has found many ways to obtain correct by design programs from high level specifiactions. One such approach \cite{itzhaky2010simple} looks to inductively synthesise invariants to generate the desired programs. Since these digital programs are already constrained by designed to factors such as FWL, these approaches are ideal for the synthesis of parametric controllers.
In \cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS for the synthesis of switching controllers for stabilizing continuous-time plants with polynomial dynamics. The work extends to its application on affine systems, finding its major challenge in the hardness of solving linear arithmetic on present SMT solvers. Since this approach uses switching states instead of linear dynamics in the digital controller, it entirely circumvents the FWL problem. It is also not suitable for the kind of control we seek to synthesise.
What is needed in this regard is the combination of a synthesis engine with a control verification tool that addresses the challenges presented here in the form of FWL effects and stability measures for LTI SISO controllers. We take the former from \cite{} and the latter from \cite{daes20161} whilst enhancing the procedure by evaluating the quantaization effects of the Hardware interfaces (ADC/DAC) to obtain an accurate discrete time FWL representation of the continuous dynamics.

>>>>

@@

First, some assumptions have to be established to obtain a model for
$\Delta{G_{Q}(z)}$.
%
\begin{myassumption}
\label{whitenoise}
%
The quantization noises $\nu_{1}$ (from the $Q1$) and $\nu_{2}$ (from the
$Q2$) are white noises with zero mean.
%
\end{myassumption}

@@

\begin{myassumption}
%
The DAC and ADC are synchronized, {\it i.e.}, there is no delay between
sampling the plant output at ADC and updating the DAC accordingly.
%
\end{myassumption}

\begin{myassumption}
%
The controller and plant are linear.  In particular, the digital controller
does not contain nonlinearities such as overflows, saturation, and
limit-cycle oscillations.
%
\end{myassumption}

\begin{myassumption}
%
The DAC interpolator is an ideal ZOH; it takes the discrete value of control
signal $u(k)$ at instant $k$ and produces a rectangular pulse of height
$u(k)$ and width equal to the sample time $T$.
%
\end{myassumption}

Given those assumptions, the present approach is further based on the following lemmas.
%
\begin{mylemma}
%
\cite{widrow1956} The second moment (variance) of the PDF of quantization
noise is given by \textcolor{red}{expand the acronym PDF}
%
\begin{equation}
\label{eq:variancelemma}
\sigma_{Q}^{2}=\frac{q^{2}}{12},
\end{equation}
\end{mylemma}

@@

\begin{mylemma} \cite{astrom1997computer}
%
The pulse transfer function $G(z)$ is equivalent at the sample times to the
continuous transfer function $G(s)$ driven by ZOH.  This pulse transfer
function can be computed as
%
\begin{equation}
\label{eq:pulsetf}
G(z)=(1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}
\end{mylemma}

@@

\noindent where $\sigma_{Q}^{2}$ is the variance of quantization noise and
$U(z)$ is the z-transform of control signal ({\it i.e.}, the z-transform of
the signal from the digital controller), which is provided to the plant by
means of ZOH.
%
\begin{proof}
%
Suppose that the plant discrete model in Figure~\ref{fig:sampledsystem} can
be represented by the following pulse transfer function (without
quantization noise effect):
%
\begin{equation}
\label{eq:tfwithout}
G(z)=\frac{Y(z)}{U(z)}=\frac{B(z)}{A(z)}=\frac{b_{0}+b_{1}z^{-1}+b_{2}z^{-2}+\cdots+b_{M}z^{-M}}{1+a_{1}z^{-1}+a_{2}z^{-2}+\cdots+a_{N}z^{-N}},
\end{equation}
%
\noindent where $Y(z)$ is the z-transform of the plant output signal without
noise $y(k)$, $U(z)$ is the z-transform of the control signal, which is
provided to the plant, $A(z)$ and $B(z)$ are the plant pulse transfer
function numerator and denominator, respectively.

@@

Eq.~\eqref{eq:deltag_var} indicates that the modeling of quantization noise
effect on the plant transfer function demands an estimation of the variance
of quantization noise $\sigma_{Q}^{2}$.  To obtain such estimation, the
methodology presented by Widrow and Koll\'ar~\cite{widrow2008quantization}
is used.  The authors provide an algorithm for computing the mean square of
the output noise due to $Q1$ and $Q2$.  Considering that each quantizer
contribution can be represented by a white noise with zero mean
(Assumption~\ref{whitenoise}), the mean square can be considered as the
variance of noise ($\sigma_{Q}^{2}$), which can be calculated as
%
\begin{equation}
\sigma_{Q}^{2}=\frac{\lambda_{1}^{2}\cdot \sum {h_{Q1}^{2}}+\lambda_{2}^{2}\cdot \sum {h_{Q2}^{2}}}{12},
\end{equation}
%
\noindent where $\lambda_{1}$ and $\lambda_{2}$ are the scale factor for the
quantization $Q1$ and $Q2$, respectively; and $\sum {h_{Q1}^{2}}$ and $\sum
{h_{Q2}^{2}}$ are the squares sums of the impulse response obtained by means
of a transfer function from $Q1$ and $Q2$ to the plant output, respectively.

@@

These equations are sufficient for computing $\sigma_{Q}^{2}$.  However,
there is still a gap for a complete model for $\Delta{G_{Q}(z)}$ in
Eq.~\eqref{eq:deltag_var}, {\it i.e.}, the z-transform of control action
$U(z)$.  This indicates that the quantization noise effect on a discrete
plant model mainly depends on three main factors:
%
\begin{itemize}
%
\item \textbf{Sample time:} the sample time effect is reflected over pulse transfer function that covers the effects of DAC ant its interpolator (ZOH);
%
\item \textbf{ADC and DAC step: } these are the main component of quantization noise power, but an $8$-bits converter already reduces the effect of quantization noises in the output for negligible values.
%
\item \textbf{Signal-to-noise ratio (SNR): } the dependence on the value of $U(z)$ shows the importance of the SNR for handling quantization noise; for a high SNR of the digital controller, the quantization effect is negligible. This happens during the transient of the system, when high control actions are employed. However, it can be relevant during the steady state. A digital controller with integral action might be sufficient for eliminating such influence on the steady state.
%
\end{itemize}

@@

\noindent where $R(z)$ can be considered as a step signal (for stabilization
purpose) with height equal to $r$, which can be rewritten as
%
\begin{equation}
\Delta{G_{Q}(z)}=\frac{\sigma^{2}_{Q}}{r}\cdot \frac{1+C(z)G(z)}{C(z)} \cdot \frac{z-1}{z}.
\end{equation}

@@

The problem of parametric control synthesis based on stability measures for
continous Linear Time Invariant (LTI) Single Input-Single Output (SISO)
systems has been researched for several decades.  On a theoretical level it
is a solved problem~\cite{wonham1967pole} for which researches continuously
seek better results based on a number of properties in addition to
stability.  A vast range of pole placement techniques such as Moore's
algorithm for eigenstructure assignment~\cite{klein1977eigenvalue} or the
more recent Linear Quadratic Regulator (LQR)~\cite{bemporad2002explicit}
have been used in several papers with increasing degrees of succes to ensure
stability.  The latter approach highlights the importance of conserving
energy during the control process, which results amongst other things in
lower running costs.  Since real systems are subject to tolerance and noise
as well as the need for economy, more recent works focus on the problems of
achieving robust stability with minimum
gain~\cite{schmid2014unified,konigorski2012pole}.  However, when extending
the problem to digital controller synthesis, many of these techniques lack
the ability to produce sound or stable results because they disregard the
effects of quantization and rounding.  Modern papers on
implementations/synthesis of LTI digital
controllers~\cite{das2013lqr,ghosh2013fpga} focus on time discretization,
failing to account for these error-inducing effects and can therefore result
in digitally unstable systems that have been proven to be robustly stable in
a continuous space.  This situation demands the inclusion of formal methods
to verify the effects of space discretization on the implementation of the
controller.

@@

Researchers in formal verification methods of control systems have studied
various effects of discretising the dynamics, including delayed
response~\cite{Duggirala2015}, and Finite Word Length (FWL) constraints with
the objective to either verify (eg.  \cite{daes20161}) or optimise
(eg.~\cite{oudjida2014design}) implementations.  There are in fact two
different problems arising from FWL effects.  The first corresponds to the
error in the dynamics caused by the inability of the hardware to represent
the exact dynamics of the system whilst the second relates to rounding
errors during computation.  In \cite{fialho1994stability} a stability
measure based on the error of the digital dynamics ensures that the FWL
errors do not make the digital system unstable.  A more recent
approach~\cite{DBLP:journals/automatica/WuLCC09} uses $\mu$ calculus to
directly model the digital controller so that the selected parameters are
stable by design.  Most papers following this line of verification focus on
finding a correct implementation of a known desired controller, looking for
optimal parameter representations using FWL but ignoring the effects of
rounding errors due to issues of mathematical tractability.  The analyses
in~\cite{DBLP:conf/hybrid/WangGRJF16,DBLP:conf/hybrid/RouxJG15} rely on an
invariant computation on the discrete system dynamics using Semi-Definite
Pogramming (SDP).  Whilst the former uses BIBO properties to determine
stability, the latter uses Lyapunov-based quadratic invariants.  In both
cases the SDP solver uses floating point arithmetic and soundness is checked
by upper bounding the error.

@@

To the authors' knowledge, there is no existing work on the automatic
synthesis of digital controllers considering FWL effects.  Parameter
synthesis tools such as~\cite{cimatti2013parameter} use bounded model
checking and fixed point computations to find parameters based on user
defined specifications, but they often operate in the continuous domain and
are ill-suited for robust analysis since they have no direct means of
evaluating robustness.  Other tools such as~\cite{economakos2016automated}
are directly aimed at robust stability problems but fail to explore the
effects of FWL in their implementation.  In order to provide a correct by
design controller, \cite{alur2016compositional} requires a user-defined
finite state abstraction to synthesise a digital controller based on high
level specifications.  Whilst this approach directly overcomes the
challenges presented by the FWL problem, it still requires error-prone user
intervention.  A different solution that looks at the other end of the scale
is an approach that synthesises word lengths for know problems as developed
in \cite{jha2013swati}, however this provides neither an optimal word
synthesis nor a comprehensive solution for the problem.  It is from this
vacuum that we seek to explore a CEGIS approach to FWL controller synthesis.

@@

In recent years the development of program synthesis has found many ways to
obtain correct by design programs from high level specifiactions.  One such
approach~\cite{itzhaky2010simple} looks to inductively synthesise invariants
to generate the desired programs.  Since these digital programs are already
constrained by designed to factors such as FWL, these approaches are ideal
for the synthesis of parametric controllers. 
In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS for the
synthesis of switching controllers for stabilizing continuous-time plants
with polynomial dynamics.  The work extends to its application on affine
systems, finding its major challenge in the hardness of solving linear
arithmetic on present SMT solvers.  Since this approach uses switching
states instead of linear dynamics in the digital controller, it entirely
circumvents the FWL problem.  It is also not suitable for the kind of
control we seek to synthesise.  What is needed in this regard is the
combination of a synthesis engine with a control verification tool that
addresses the challenges presented here in the form of FWL effects and
stability measures for LTI SISO controllers.  We take the former
from~\cite{} and the latter from \cite{daes20161} whilst enhancing the
procedure by evaluating the quantaization effects of the Hardware interfaces
(ADC/DAC) to obtain an accurate discrete time FWL representation of the
continuous dynamics.
